---
title: "Danny fitness 2015"
author: "Martijn Vellinger"
date: "28 Dec 2015"
output: html_document
---
```{r, preload}
library(dplyr)
library(ggplot2)
library(scales)
library(lubridate)
library(reshape2)
dark_bar_theme <- function() {
    theme(
        axis.text.x = element_text(angle = 90, size = 7, family = "Montserrat", colour = "grey90", hjust = 1),

        axis.text.y = element_text(size = 7, family = "Montserrat", colour = "grey90", hjust = 1),

        plot.title = element_text(colour = "grey90", size = 12, family = "Montserrat", hjust = 1),

        axis.title = element_text(colour = "grey90", size = 8, family = "Montserrat", hjust = 0),

        plot.background = element_rect(fill = "grey20", colour = "grey25"),
        panel.background = element_rect(fill = "grey20"),
        panel.grid.major.x = element_line(colour = "grey80", linetype = "dotted", size = 0.2),
        panel.grid.minor.x = element_line(colour = "grey60", linetype = "dotted", size = 0.2),
        panel.grid.major.y = element_line(colour = "grey80", linetype = "dotted", size = 0.2),
        panel.grid.minor.y = element_line(colour = "grey60", linetype = "dotted", size = 0.2),
        panel.border = element_blank(),
        legend.justification=c(1,1),
        legend.position=c(1,1),
        legend.background = element_blank(),
        legend.title = element_text(size = 12,
                                    family = "Montserrat",
                                    colour = "grey90",
                                    hjust = 1),
        legend.text = element_text(size = 9,
                                   family = "Montserrat",
                                   colour = "grey90",
                                   hjust = 1),
        legend.key = element_blank(),
        legend.key.size = unit(20, "native")
    )
}
```

# Reading the data

Danny has supplied 3 export files from his various fitness tracking tools (FitNotes, Libra and MyFitnessPal).
Our first order of business is loading these in and looking what's there. Annoyingly, the .csv files do not use the same separator character, so we'll specify each:

```{r, read.data}
fitnotes <- read.csv("FitNotes_Export.csv", sep=",", stringsAsFactors = F)
libra <- read.csv("Libra_2015-12-28.csv", sep=";", stringsAsFactors = F)
myfitnesspal <- read.csv("MyFitnessPal_Food_Data_Export.csv", sep=",", stringsAsFactors = F)

head(fitnotes, n=3)
head(libra, n=3)
head(myfitnesspal, n=3)
```

Obviously there are a few issues here, first of which is that the date fields are different in each file.
That is going to be the first thing we fix by adding a new column to each data set with a proper date object.

```{r, fix.dates}
myfitnesspal$p.date <- as.Date(strptime(myfitnesspal$Date, "%d-%b-%y"))
libra$p.date <- as.Date(strptime(libra$X.date, "%d/%m/%y %H:%M"))
fitnotes$p.date <- as.Date(fitnotes$Date)
```

Not all of the data is useful, so we're going to reduce our data frames and keep only what we want using dplyr:

```{r, reduce.data}
myfitnesspal <- select(myfitnesspal, Calories, Carbs, Fat, Protein, Cholesterol, Sodium, Sugars, Fibre, p.date)

fitnotes <- select(fitnotes, Exercise, Category, Weight..kgs., Reps, p.date)

libra <- select(libra, weight, p.date)

# Sodium is in mg. everything else in g. so we're going to correct that:
myfitnesspal$Sodium <- myfitnesspal$Sodium / 1000 
myfitnesspal$Cholesterol <- myfitnesspal$Cholesterol / 1000
```

Let's see how Danny's meal composition changes over time:

```{r, melt}
macros <- select(myfitnesspal, p.date, Protein, Carbs, Fat)
macros$week <- week(macros$p.date)
macros.agg <- arrange(aggregate(.~week, macros, FUN =mean), by=week)

calories <- select(myfitnesspal, p.date, Calories) 
calories$week <- week(calories$p.date)
calories.agg <- aggregate(.~week, calories[2:3], FUN = mean)
macros.melt <- melt(macros.agg, id.vars="week", measure.vars = c("Protein", "Carbs", "Fat") ) #create a long form data frame
calories.melt <- melt(calories.agg, id.vars="week") #create a long form data frame

libra.subset <- subset(libra, p.date < max(calories$p.date + 1) & p.date > min(calories$p.date -1))


#normalize low calorie measurements (meals which were not logged)
calories$rollmean <- rollmean(calories$Calories, k=3)[1:nrow(calories)]

calories$Calories[calories$Calories < 1600] <- calories$rollmean[calories$Calories < 1600]

calories <- select(calories, p.date, Calories)
#mergedcvsw$Calories[mergedcvsw$Calories < 1800] <- mean(mergedcvsw$Calories[mergedcvsw$Calories > 1600])
#mergedcvsw$Calories[mergedcvsw$Calories > 3200] <- mean(mergedcvsw$Calories[mergedcvsw$Calories > 1600])
#normalize calorie values down to a more reasonable range:

mergedcvsw <- merge(calories, libra.subset, by = "p.date")

merged.melt <- melt(select(mergedcvsw, p.date, Calories, weight), id.vars = "p.date")

ggplot(data = merged.melt, aes(x=p.date, y=value, fill=variable, colour=variable)) + geom_point(size=3) + geom_point(size=1, colour="grey30") + geom_line(stat="identity") + geom_smooth() + dark_bar_theme() + scale_x_date(breaks = "week") + facet_grid(variable~., scales="free_y") + scale_color_manual(values = c("#bdffd8", "#fcff66"))


#macros
ggplot(data=macros.melt, aes(x = week, y = value, fill = variable)) + geom_bar(stat="identity", position = position_dodge()) + dark_bar_theme() + scale_x_continuous(breaks = seq(1,51,1)) + scale_y_continuous() + scale_fill_manual(values = c("#adf0ca", "#73acba", "#ffff82"), name="VARIABLE")

#calories
ggplot(data = merged.melt, aes(x=p.date, y=value, fill=variable, colour=variable)) + geom_point(size=3) + geom_point(size=1, colour="grey30") + geom_line(stat="identity") + dark_bar_theme() + scale_x_date(breaks = "week")


